---
title: "Sample size calculation for future trials"
author: "Akihiro Shiroshita"
date: "2021/2/8"
output: pdf_document
---

```{r setup, include=FALSE}
knitr::opts_chunk$set(echo = TRUE)
```

## Sample size calculation  
```{r include=FALSE}
library(tidyverse)
library(meta)
library(Hmisc)
```

### Meta-analysis  
Nanni 2018, Pimentel 2019, Zhao 2017  

```{r cars, echo=TRUE}
df <- read.csv("data.csv", sep="," , header=T)
metaprop(event=Event,
         n=Sample,
         data=df,
         studlab=Study,
         sm="Plogit",
         comb.fixed=FALSE,
         comb.random=TRUE,
         hakn=F,
         method.tau="DL")
```
  
The pooled 3-year mortality was calculated.    

## Sample size calculation  
```{r pressure, echo=TRUE}
tref <- 3 #time at which mortalities estimated
Nsim <- seq(5000,6000, 100) 
mc <- 0.74 #tref-year mortality, control
hr <- 1.13 #hazard ratio.
r <- (1-((1-(1-mc)^hr)/mc))*100 #% reduction in mc by intervention
accrual <- 2 #duration of accrual period
tmin <- 2/12 #minimum follow-up time
pwr.res<-c()
for (i in 1:length(Nsim) ){
  pwr.tmp<-cpower(tref=tref,
                  n=Nsim[i],
                  mc=mc,
                  r=r,
                  accrual=accrual,
                  tmin=tmin,
                  noncomp.c=0,
                  noncomp.i=0,
                  alpha=0.05,
                  pr=FALSE)
  pwr.res[i]<-pwr.tmp["Power"]
}
pwr.res.all<-cbind(Nsim, pwr.res)
pwr.res.all
``` 
  
A total of 5200 patients will be needed.  

## Methods  
*Sample size calculation for future trials*  
We performed sample size calculation for future trials targeting at patients with metastatic/recurrent breast cancer receiving chemotherapy or endocrine therapy. First, we picked up the median follow-up period among the included studies and calculated the pooled event rates using the random-effects model. Then, we simulated trials of comparing the entire survival curves between metformin group and placebo group. We used the following data; $\alpha = 0.05$, $power = 0.8$, study accrual period, percent of drop-ins and non-adherers, the pooled event rates of placebo during the median follow-up period, and the pooled hazard ratio. Parameters were based on the results of systematic review. R version 3.6.0 (R core Team, Vienna, Austria) and packages "meta" (version. 4.16-2) and "Hmisc" (version 4.4-0) were used for sample size calculation.  

## Results  
*Sample size calculation for future trials*  
Based on the information gathered from our systematic review, we set the parameters as following; study accrual period of 1 year, no drop-ins or non-adherers, the pooled event rates of 0.74 among the placebo group during the 3-year follow-up period, and the pooled hazard ratio of 1.13.  When randomised controlled trials with a allocation ratio of 1:1 are conducted, a total of 5200 patients will be needed to detect the difference assuming $\alpha = 0.05$ and $power = 0.8$.